---
title: "fPAR Modelling"
author: "Jeff Atkins"
date: "February 17, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modelling fPAR in respect to CSC metrics (i.e. rugosity and porosity)

I have subset the data into direct and diffuse light regimes. The a priori assumption here is based on Beer's Law.

```{r data_read, echo = FALSE}
master <- read.csv("master_light_csc.csv")

diffuse <- subset(master, master$regime == "diffuse")
direct <- subset(master, master$regime == "direct")
```

## Direct Light - fPAR as a function of Rugosity

First, approximating fPAR under direct light using the best model fit:

$fPAR = (a * R_{c}) / (b + R_{c})$

```{r direct, echo = FALSE}
y <- direct$fPAR
x <- direct$rugosity


m.direct <- nls(y ~ (a * x) / (b + x), start = list(a = 1, b = 1))
summary(m.direct)

RSS <- sum(residuals(m.direct)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)
message("R squared = ")
print(R.square)

x11()
plot(x, y, xlab = "Rugosity", ylab = "fPAR", main = "DIRECT LIGHT")
a <- coef(m.direct)[1]
b <- coef(m.direct)[2]
curve((a * x) / (b + x), col = "blue", add = TRUE)
```

## Direct Light - fPAR as a function of Porosity

Some model exploration seems to point towards a third-order polynomial being a first-order approximation for model fit. (Can refine later):

```{r porosity, echo = FALSE}
x <- direct$porosity
y <- direct$fPAR

rm(a)
rm(b)
m.3 <- lm(y ~ poly(x, 3))
summary(m.3)

m.poly <- function(x) m.3$coefficient[4]

x11()
plot(x,y, xlab = "porosity", ylab = "fPAR", main = "fPAR and Porosity--direct light")
lines(sort(x), fitted(m.3)[order(x)], col = "blue")
```


## Residuals analysis one

If we extract the residuals from our fPAR ~ $R_c$ model against canopy porosity, it is mostly noise:

```{r rugosity_resid, echo = FALSE}
y <- residuals(m.direct)
x <- direct$porosity

x11()
plot(x, y, xlab = "Porosity", ylab = "residuals (fPAR ~ Rc model)", main = "Residuals of fPAR ~ Rc against Porosity")
```

Now extracting residuals from our fPAR ~ Porosity model against canopy rugosity:

```{r porosity_resid, echo = FALSE}
# residuals of the fPAR ~ porosity model against rugosity
# 

y <- residuals(m.3)
x <- direct$rugosity

x11()
plot(x, y, xlab = "Rugosity", ylab = "residuals (fPAR ~ Pc model)", main = "Residuals of fPAR ~ Pc against Rugosity")
```
