---
title: "JeffPar Model fitting"
author: "Jeff Atkins"
date: "July 14, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r data-import, echo = FALSE, warning=FALSE, message=FALSE}
#
require(plyr)
require(ggplot2)
require(dplyr)
require(vegan)
require(ggrepel)
require(magrittr)
require(nlstools)

#adding in palette
library(RColorBrewer)
darkcols <- brewer.pal(10, "Dark2")

#Data import
plot.fpar <- read.csv("laserquest_master_fpar_plot.csv")

# designate light regime based on aPAR or above canopy PAR
plot.fpar$regime <- ifelse(plot.fpar$aPAR_mean >= 1000, "high",
                             "low")

plot.fpar$regime <- as.factor(plot.fpar$regime)

### Modeling section
#Subset data for modelling
high <- subset(plot.fpar, aPAR_mean >= 1000)
low <- subset(plot.fpar, aPAR_mean < 1000)

#new.master <- read.csv("laserquest_plot_means_with_fPAR.CSV")
#custom plot theme for all
theme_new <- function(base_size = 12){
     theme_bw(base_size = base_size) %+replace%
          theme(
               #line = element_line(colour="black"),
               #text = element_text(colour="black"),
               #axis.text = element_text(colour="black", size=8),
               #strip.text = element_text(size=12),
               #legend.key=element_rect(colour=NA, fill =NA),
               panel.grid = element_blank(),   
               # panel.border = element_blank(),
               panel.background = element_rect(fill = "white", colour = "white"), 
               strip.background = element_rect(fill = NA),
               axis.text = element_text( size = 14),
               axis.ticks.x = element_blank(),
               axis.ticks.y = element_blank(),
               axis.title  = element_text( size = 16, margin = margin(12, unit = "cm"))
          )
     
}
# custom plot label

vai.label =  expression(paste(VAI~(m^2 ~m^-2)))
lai.label =  expression(paste(LAI~(m^2 ~m^-2)))
log.lai.label = expression(paste(log~( LAI~(m^2 ~m^-2))))
r.label = expression("R"[C])
fpar.label = expression(paste(italic("f"), "PAR"))

```

## Rectangular Hyperbola Option 1
Focusing on model fit for plot means for fPAR and Rumple under high light conditions.
First with the equation $y = a*x/b+x$


```{r one1, echo = FALSE, messages = FALSE}

# RUMPLE
#HIGH
y <- high$fPAR_mean
x <- high$rumple_mean


high.rumple <- nls(y ~ (a * x) / (b + x), start = list(a = 1, b = 1))
#summary(high.rumple)
message('Model Coefficients')
coef(high.rumple)
message("AIC")
AIC(high.rumple)


RSS <- sum(residuals(high.rumple)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)

R.square <- paste("R^2 == ", round(R.square, 3))
# 
# message('R squared = ')
# print(R.square)

fpar.high.rumple <- function(x) {(2.420 * x) / (10.672 + x) }
# fpar.high.rumple2 <- function(x) {2.849 + (-1.660 * x) / ( -0.821 + x)}
# fpar.high.rumpl3 <- function(x) { 1 / (1 + exp(-0.7972 * (x - 2.5677)))}
ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)

```
\newpage

## Rectangular Hyperbola Option 2

Using the equation $y = a + b*x/c+x$

```{r two, echo = FALSE, messages = FALSE}
high.rumple2 <- nls(y ~ a + (b * x) / (c + x), start = list(a = 0, b = 0.1, c = 0.1))
#summary(high.rumple2)
message("Model Coefficients")
coef(high.rumple2)
message("AIC")
AIC(high.rumple2)

RSS <- sum(residuals(high.rumple2)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)

R.square <- paste("R^2 == ", round(R.square, 3))
# 
# 
# message("R squared =")
# print(R.square)
a <- coef(high.rumple2)[1]
b <- coef(high.rumple2)[2]
c <- coef(high.rumple2)[3]

fpar.high.rumple2 <- function(x) {a + (b * x) / ( c + x)}
ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple2, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple2)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)
```
\newpage

## Sigmoidal Function

Using the equation:  $y = 1/(1 + exp(a * (x - b)))$

```{r pressure, echo=FALSE, messages = FALSE}
high.rumple3 <- nls(y ~ (1 / (1 + exp(-a * (x - b)))), start = list(a = 1, b = 1))
message("Model Coefficients")
coef(high.rumple3)
message("AIC")
AIC(high.rumple3)


RSS <- sum(residuals(high.rumple3)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)


R.square <- paste("R^2 == ", round(R.square, 3))


a <- coef(high.rumple3)[1]
b <- coef(high.rumple3)[2]

fpar.high.rumple3 <- function(x) {1 / (1 + exp(-a * (x - b)))}

ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple3, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple3)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)
```

\newpage

## Logistic Growth Model

Using the equation:  $y = a / (1 + b * exp(-k * x)$

```{r four, echo=FALSE, messages = FALSE}
high.rumple4 <- nls(y ~ a / (1 + b * exp(-k * x)), start = list(a = 1, b = 1, k = 1))
message("Model Coefficients")
coef(high.rumple4)
message("AIC")
AIC(high.rumple4)


RSS <- sum(residuals(high.rumple4)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)
#R.square <- as.character(round(R.square, 3)

R.square <- paste("R^2 == ", round(R.square, 3))
# message("R squared =")
# print(R.square)

a <- coef(high.rumple4)[1]
b <- coef(high.rumple4)[2]
k <- coef(high.rumple4)[3]

fpar.high.rumple4 <- function(x) {a / (1 + b * exp(-k * x))}

ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple4, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple4)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)
```


\newpage

## Logarithmic Model

Using the equation:  $y = a + b ln(x)$

```{r five, echo=FALSE, messages = FALSE}
high.rumple5 <- nls(y ~ a + b * log(x), start = list(a = 1, b = 1))
message("Model Coefficients")
coef(high.rumple5)
message("AIC")
AIC(high.rumple5)


RSS <- sum(residuals(high.rumple5)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)
#R.square <- as.character(round(R.square, 3)

R.square <- paste("R^2 == ", round(R.square, 3))
# message("R squared =")
# print(R.square)

a <- coef(high.rumple5)[1]
b <- coef(high.rumple5)[2]


fpar.high.rumple5 <- function(x) {a + b * log(x)}

ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple5, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple5)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)
```

\newpage

## Rise to Max, Raise the Roof

Using the equation:  $y = a * (1 - exp(-b * x)) + c * (1 - exp(-d * x))$

```{r six, echo=FALSE, messages = FALSE}

high.rumple6 <- nls(y ~ (a * (1 - exp(b * x))) + (c * (1 - exp(d * x))), start = c(a = 1, b = 1, c = 1, d = -1), control = nls.control(warnOnly = TRUE))
message("Model Coefficients")
coef(high.rumple6)
message("AIC")
AIC(high.rumple6)


RSS <- sum(residuals(high.rumple6)^2)
TSS <- sum((y - mean(y))^2)
R.square <- 1 - (RSS/TSS)
#R.square <- as.character(round(R.square, 3)

R.square <- paste("R^2 == ", round(R.square, 3))
# message("R squared =")
# print(R.square)

a <- coef(high.rumple6)[1]
b <- coef(high.rumple6)[2]
c <- coef(high.rumple6)[3]
d <- coef(high.rumple6)[4]


fpar.high.rumple6 <- function(x) {a * (1 - exp(b * x)) + c * (1 - exp(d * x))}

ggplot(high, aes(x = rumple_mean, y = fPAR_mean))+
     geom_point(size = 3, shape = 21)+
     theme_new()+
     xlim(c(0,15))+
     scale_y_continuous(breaks = seq(0, 1, 0.5), limits = c(0, 1))+
     xlab("Rumple")+
     ylab(fpar.label)+
     stat_function(fun = fpar.high.rumple6, col = "blue", size = 1)+
     annotate("text", x = 10, y = 0.4, label =  R.square, parse = TRUE)

nr <- nlsResiduals(high.rumple6)
plot(nr, which = 0)
plot(nr, which = 5)
plot(nr, which = 6)
```