---
title: "R test run of rugosity code"
author: "Jeff Atkins"
date: "October 5, 2016"
output: pdf_document
---
```{r setup1, echo = FALSE }
require(ggplot2)

```
## OVERVIEW

This document runs through the testing of a draft module of an R package to be used to calculate canopy structural metrics from portable canopy LiDAR (PCL) data. This code is meant to reproduce methods as outlined in Hardiman et al. 2011 etc.

The code now pulls from an functions.R script that includes the sub-functions that parse and process the PCL data files. 


```{r setup, echo = TRUE }
# Source functions
source("functions.R")

```

## WORKFLOW

The draft of the script runs a single-transect at a time and requires the entry of a data directory (noted as: `data_dir`) and the name of the file to be imported (noted as: `filename`). The script then uses the function `read.pcl` that uses a custom `read.csv` format to import and structure PCL data. A diagnostic plot of the data can also be produced with the function `pcl.diagnostic.plot`.

```{r import.files, echo = TRUE}
data_dir <- "./data/rice/"
filename <- "rice_control_two.CSV"

#function to import data
test.data <- read.pcl(data_dir, filename)

# let's take a quick look at the data
head(test.data)

pcl.diagnostic.plot(test.data, "RICE Control 2", -1e+08)

```

The next step includes adjusting by user height with the function `adjust_by_user` that takes the data frame name and the "Brady Height" value (the distance of the PCL from the ground in meters). LiDAR "hits" are then coded using `code_hits`--a function that allows for processing the data through vectorization. The data are then binned to x and z bins using the function `split_transects_from_pcl`.


```{r making.bins, echo = TRUE}
test.data <- adjust_by_user(test.data, 1.2)

test.data <- code_hits(test.data)

#the bread and butter of this code
test.data.binned <- split_transects_from_pcl(test.data, 30, 10)

#let's look at the data again
head(test.data.binned)
summary(test.data.binned)
```

Next comes the formulation of the matrix for the basis of further calculations. The matrix results from condensing data into a matrix with dimensions of length of transect X maximum measured canopy height. For example, a 40 m transect with max canopy height of 29.3 m would results in a matrix with dimensions of 40 x 29. The function `make_matrix` that does this gives some extra info as well to be cleaned up.
```{r make.matrix, echo = TRUE}
m <- make_matrix(test.data.binned)
summary(m)
```
Now we calculate VAI using the following function `calc_vai` which is included for demo purposes:

```{r demo.function, echo = TRUE}
calc_vai <- function(df) {
     
     df$vai <- (df$bin.hits / df$lidar.pulses) 
     df$vai <- df$vai * 8 #adjust for max lai?
     df$vai <- df$vai * -1
     df$vai <- log(1.0 - df$vai*0.9817)/0.5
     df[is.na(df)] <- 0
     return(df)
 
}
```
```{r vai, echo = TRUE}
m <- calc_vai(m)
```

Then comes Rugosity, but it is the end of the day and I need to get home so here is the rest which also includes a draft hit grid:

```{r restofcode, echo = TRUE}
#######new test
m <- calc_mean_leaf_ht(m)
m <- calc_std_bin(m)

#the jam
calc_rugosity(m)

#hit grid
ggplot(m, aes(x = xbin, y = zbin))+ 
     geom_tile(aes(fill = vai))+
     scale_fill_gradient(low="palegreen1", high="dark green", 
                         name="LiDAR\n Method One")+
     #scale_y_continuous(breaks = seq(0, 20, 5))+
     # scale_x_continuous(minor_breaks = seq(0, 40, 1))+
     theme(axis.line = element_line(colour = "black"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           panel.background = element_blank())+
     xlim(0,40)+
     ylim(0,40)+
     xlab("Distance along transect (m)")+
     ylab("Height above ground (m)")




```


