---
title: "VAI"
author: "Jeff Atkins"
date: "September 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Looking at VAI

I wanted to look at the VAI calculations and walk-through what they are doing to make sure I understand them.

I am going to do this with some ordered sequences--assuming that max LAI is 8. Graphing this distribution of points along the index they occur gives us a straight line.

```{r vai.lonely}
vai.seq <- seq(0.1, 8, by =0.01)

test.vai <- data.frame(vai.seq)
plot(test.vai$vai.seq, xlab = "Index", ylab = "VAI")
```



## Now to apply the extinction equation

This produces a curve if we just adjust with only the negative log function.

```{r vai.extinct}
test.vai$vai.seq <- vai.seq * -1


test.vai$vai.seq <- log(1.0 - test.vai$vai.seq*0.9817)/0.5
plot(test.vai$vai.seq, xlab = "Index", ylab = "VAI")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
